Exact asymptotic expressions for local clustering in hyperbolic random graphs

### 2018-10-18 clustering_infinite_model.tex
updated figure 3, adjusted computation lemma 3.2 and added text on scaling k_n

\end{tikzpicture}
\caption{Example configuration of two points $p$ and $p^\prime$ for which $\BallPon{p} \cap \BallPon{p^\prime}$ is not a subset of $\BallPo{p} \cap \BallPo{p^\prime}$. The red region indicates the area of points belonging to $\BallPon{p} \cap \BallPon{p^\prime}$ but not to $\BallPo{p} \cap \BallPo{p^\prime}$.}
\end{figure}

@@

\end{tikzpicture}
\caption{Situation for the intersections of the connection intervals considered in Lemma~\ref{lem:ordered}, with $y_0 < y_1 <y_2$ and $0 \le x_1 \le e^{(y_0 + y_1)/2}$. The regime for $x_2$ in which case a triangle is formed is colored in red on both intervals. \PvdH{Maybe show the full connection balls.}}
\label{fig:triangle_prob_lemma}
\end{figure}

>>>>

\end{tikzpicture}
\caption{Situation for the intersections of the connection intervals considered in Lemma~\ref{lem:ordered}, with $y_0 < y_1 <y_2$ fixed and for different cases of $0 \le x_1 \le e^{(y_0 + y_1)/2}$. The top figure shows the case where $0 \le x_1 \le e^{(y_1 + y_2)/2} - e^{(y_0 + y_2)/2}$, while the bottom one shows the case $x_1 > e^{(y_1 + y_2)/2} - e^{(y_0 + y_2)/2}$. The solid red line indicates the range for $x_2$ such that the points $p_0$, $p_1$ and $p_2$ form a triangle. The boundaries of there balls are show in, respectively, black, blue and green.}
\label{fig:triangle_prob_lemma}
\end{figure}

@@

\end{tikzpicture}
\end{figure}

### 2018-10-18 concentration_argument.tex
updated figure 3, adjusted computation lemma 3.2 and added text on scaling k_n

\begin{lemma}\label{lem:average_degree_hyperbolic}
Let $\eps \in (0, 1)$. Then for all $0 \le y \le (1 - \eps)R_n$
\[
	 1 - \phi_{\H,n}^{(1)}(y) - \phi_{\H,n}^{(2)}(y) \le \frac{\Mu{\BallHyp{0,y}}}{\Mu{\BallPo{0,y}}} 
	 \le 1 - \phi_{\H,n}^{(1)}(y) + \phi_{\H,n}^{(2)}(y),
\]
where
\[
	\phi_{\H,n}^{(1)}(y) = \frac{2\alpha - 3}{2} e^{-(\alpha - \frac{1}{2})(R_n - y)} 
		+ \frac{2\alpha - 1}{2}e^{-(\alpha - \frac{1}{2})R_n - y/2},
\]
and
\[
	\phi_{\H,n}^{(2)}(y) = 
	+ \begin{cases}
			\frac{(2\alpha - 1)K}{3 - 2\alpha}\left(e^{-(\alpha - \frac{1}{2})(R_n - y)} - e^{-(R_n - y)}\right)
			&\mbox{if } 1/2 < \alpha < 3/2,\\
			\frac{(2\alpha -1)K}{2} (R_n - y)e^{-(R_n - y)} &\mbox{if } \alpha = 3/2,\\
			\frac{(2\alpha - 1)K}{2\alpha - 3} \left(e^{-(R_n - y)} - e^{-(\alpha - \frac{1}{2})(R_n - y)}\right)
			&\mbox{if } \alpha > 3/2.
		\end{cases}
\]
\end{lemma}

\begin{proof}
We will split the computation of $\Mu{\BallHyp{0,y}}$ over the case $y^\prime > R_n - y$ and $y^\prime \le R_n - y$ where for the latter we utilize Lemma~\ref{lem:asymptotics_Omega_hyperbolic}. Recall that $\Mu{\BallPo{0,y}} = \xi_{\alpha,\nu} e^{y/2}$ where $\xi_{\alpha,\nu} = 4\nu \alpha/(2\alpha - 1)\pi$. \PvdH{Check computations below for new $\xi_{\alpha,\nu} = 4\nu \alpha/(2\alpha - 1)\pi$}

By~\eqref{eq:tail_inclusion_hyperbolic_ball}, we have $\BallHyp{(0,y)} \cap \Rcal_n ([R_n - y, R_n])= \Rcal_n([R_n-y,R_n])$. 
Thus, 
\begin{align*}
	&\hspace{-30pt}\Mu{\BallHyp{(0,y)} \cap \Rcal_n ([R_n - y, R_n])} \\
	&= \int_{R_n - y}^{R_n} \int_{I_n} f_{\alpha,\nu}(x^\prime, y^\prime) \dd x^\prime \dd y^\prime
		= \nu \alpha e^{R_n/2}\left(e^{-\alpha(R_n - y)} - e^{-\alpha R_n}\right)\\
	&= \Mu{\BallPo{0,y}} \frac{2\alpha - 1}{2}\left( e^{-(\alpha - \frac{1}{2})(R_n - y)}
		- e^{-(\alpha - \frac{1}{2})R_n - y/2}\right) \numberthis \label{eq:mu_hyperbolic_ball_part_1}
\end{align*}

Next we will establish upper and lower bounds on $\Mu{\BallHyp{(0,y)} \cap\Rcal_n[(0,R_n-y)]}$. Using Lemma~\ref{lem:asymptotics_Omega_hyperbolic} we have 
\begin{align*}
	\Mu{\BallHyp{(0,y)} \cap\Rcal_n[(0,R_n-y)]} 
	&\le \nu \alpha \int_{0}^{R_n - y} \left(e^{\frac{y + y^\prime}{2}} + K e^{\frac{3}{2}(y + y^\prime) - R_n}\right)
		e^{-\alpha y^\prime} \dd y^\prime\\
	&= \Mu{\BallPo{0,y}}\left(1 - e^{-(\alpha - \frac{1}{2})(R_n - y)}\right) \\
	&\hspace{10pt}+ \nu \alpha K e^{\frac{3 y}{2} - R_n}\int_0^{R_n-y} e^{(\frac{3}{2} - \alpha)y^\prime} \dd y^\prime
\end{align*}
The last integral depend on the value of $\alpha$,
\[
	\int_0^{R_n-y} e^{(\frac{3}{2} - \alpha)y^\prime} \dd y^\prime
	= \begin{cases}
		\frac{2}{3 - 2\alpha}\left(e^{(\frac{3}{2} - \alpha)(R_n - y)} - 1\right) &\mbox{if } 1/2 < \alpha < 3/2,\\
		R_n - y &\mbox{if } \alpha = 3/2,\\
		\frac{2}{2\alpha-3}\left(1 - e^{-(\alpha - \frac{3}{2})(R_n - y)}\right) &\mbox{if } \alpha > 3/2.
	\end{cases}
\]
Therefore we get
\begin{align*}
	&\hspace{-30pt}\nu \alpha K e^{\frac{3 y}{2} - R_n}\int_0^{R_n-y} e^{(\frac{3}{2} - \alpha)y^\prime} \dd y^\prime\\
	&= \Mu{\BallPo{0,y}}\begin{cases}
		\frac{(2\alpha - 1)K}{3 - 2\alpha}\left(e^{-(\alpha - \frac{1}{2})(R_n - y)} - e^{-(R_n - y)}\right)
		&\mbox{if } 1/2 < \alpha < 3/2,\\
		\frac{(2\alpha -1)K}{2} (R_n - y)e^{-(R_n - y)} &\mbox{if } \alpha = 3/2,\\
		\frac{(2\alpha - 1)K}{2\alpha - 3} \left(e^{-(R_n - y)} - e^{-(\alpha - \frac{1}{2})(R_n - y)}\right)
		&\mbox{if } \alpha > 3/2.
	\end{cases}\\
	&= \Mu{\BallPo{0,y}} \phi_{\H,n}^{(2)}(y)
\end{align*}
and hence
\[
	\Mu{\BallHyp{(0,y)} \cap\Rcal_n[(0,R_n-y)]} = \Mu{\BallPo{0,y}}\left(1 - e^{-(\alpha - \frac{1}{2})(R_n - y)} 
	+ \phi_{\H,n}^{(2)}(y)\right).
\]
Combining this with~\eqref{eq:mu_hyperbolic_ball_part_1} yields the required upper bound.

The lower bound follows by observing that the only difference with the above computations is the change of sign in front of 
\[
	\nu \alpha K e^{\frac{3 y}{2} - R_n}\int_0^{R_n-y} e^{(\frac{3}{2} - \alpha)y^\prime} \dd y^\prime.
\]
\end{proof}

>>>>

\begin{lemma}\label{lem:average_degree_hyperbolic}
Let $\eps \in (0, 1)$. Then for all $0 \le y \le (1 - \eps)R_n$
\[
	 1 - \phi_{\H,n}^{(1)}(y) - \phi_{\H,n}^{(2)}(y) \le \frac{\Mu{\BallHyp{0,y}}}{\Mu{\BallPo{0,y}}} 
	 \le 1 - \phi_{\H,n}^{(1)}(y) + \phi_{\H,n}^{(2)}(y),
\]
where
\[
	\phi_{\H,n}^{(1)}(y) = \frac{2\alpha - 1 - 4\pi}{4\pi} e^{-(\alpha - \frac{1}{2})(R_n - y)} 
		+ \left(\alpha - \frac{1}{2}\right)\pi e^{-(\alpha - \frac{1}{2})R_n - y/2},
\]
and
\[
	\phi_{\H,n}^{(2)}(y) = 
	+ \begin{cases}
			\frac{(2\alpha - 1)K}{3 - 2\alpha}\left(e^{-(\alpha - \frac{1}{2})(R_n - y)} - e^{-(R_n - y)}\right)
			&\mbox{if } 1/2 < \alpha < 3/2,\\
			\frac{(2\alpha -1)K}{2} (R_n - y)e^{-(R_n - y)} &\mbox{if } \alpha = 3/2,\\
			\frac{(2\alpha - 1)K}{2\alpha - 3} \left(e^{-(R_n - y)} - e^{-(\alpha - \frac{1}{2})(R_n - y)}\right)
			&\mbox{if } \alpha > 3/2.
		\end{cases}
\]
\end{lemma}

\begin{proof}
We will split the computation of $\Mu{\BallHyp{0,y}}$ over the case $y^\prime > R_n - y$ and $y^\prime \le R_n - y$ where for the latter we utilize Lemma~\ref{lem:asymptotics_Omega_hyperbolic}. Recall that $\Mu{\BallPo{0,y}} = \xi_{\alpha,\nu} e^{y/2}$ where $\xi_{\alpha,\nu} = 4\nu \alpha/(2\alpha - 1)\pi$.

By~\eqref{eq:tail_inclusion_hyperbolic_ball}, we have $\BallHyp{(0,y)} \cap \Rcal_n ([R_n - y, R_n])= \Rcal_n([R_n-y,R_n])$. 
Thus, 
\begin{align*}
	&\hspace{-30pt}\Mu{\BallHyp{(0,y)} \cap \Rcal_n ([R_n - y, R_n])} \\
	&= \int_{R_n - y}^{R_n} \int_{I_n} f_{\alpha,\nu}(x^\prime, y^\prime) \dd x^\prime \dd y^\prime
		= \nu \alpha e^{R_n/2}\left(e^{-\alpha(R_n - y)} - e^{-\alpha R_n}\right)\\
	&= \Mu{\BallPo{0,y}} \frac{2\alpha - 1}{4\pi} \left( e^{-(\alpha - \frac{1}{2})(R_n - y)}
		- e^{-(\alpha - \frac{1}{2})R_n - y/2}\right) \numberthis \label{eq:mu_hyperbolic_ball_part_1}
\end{align*}

Next we will establish upper and lower bounds on $\Mu{\BallHyp{(0,y)} \cap\Rcal_n[(0,R_n-y)]}$. Using Lemma~\ref{lem:asymptotics_Omega_hyperbolic} we have 
\begin{align*}
	\Mu{\BallHyp{(0,y)} \cap\Rcal_n[(0,R_n-y)]} 
	&\le \frac{2\nu \alpha}{\pi} \int_{0}^{R_n - y} \left(e^{\frac{y + y^\prime}{2}} + K e^{\frac{3}{2}(y + y^\prime) - R_n}\right)
		e^{-\alpha y^\prime} \dd y^\prime\\
	&= \Mu{\BallPo{0,y}}\left(1 - e^{-(\alpha - \frac{1}{2})(R_n - y)}\right) \\
	&\hspace{10pt}+ \frac{2\nu \alpha}{\pi} K e^{\frac{3 y}{2} - R_n}\int_0^{R_n-y} e^{(\frac{3}{2} - \alpha)y^\prime} \dd y^\prime
\end{align*}
The last integral depend on the value of $\alpha$,
\[
	\int_0^{R_n-y} e^{(\frac{3}{2} - \alpha)y^\prime} \dd y^\prime
	= \begin{cases}
		\frac{2}{3 - 2\alpha}\left(e^{(\frac{3}{2} - \alpha)(R_n - y)} - 1\right) &\mbox{if } 1/2 < \alpha < 3/2,\\
		R_n - y &\mbox{if } \alpha = 3/2,\\
		\frac{2}{2\alpha-3}\left(1 - e^{-(\alpha - \frac{3}{2})(R_n - y)}\right) &\mbox{if } \alpha > 3/2.
	\end{cases}
\]
Therefore we get
\begin{align*}
	&\hspace{-30pt}\frac{2\nu \alpha}{\pi} K e^{\frac{3 y}{2} - R_n}
		\int_0^{R_n-y} e^{(\frac{3}{2} - \alpha)y^\prime} \dd y^\prime\\
	&= \Mu{\BallPo{0,y}}\begin{cases}
		\frac{(2\alpha - 1)K}{3 - 2\alpha}\left(e^{-(\alpha - \frac{1}{2})(R_n - y)} - e^{-(R_n - y)}\right)
		&\mbox{if } 1/2 < \alpha < 3/2,\\
		\frac{(2\alpha -1)K}{2} (R_n - y)e^{-(R_n - y)} &\mbox{if } \alpha = 3/2,\\
		\frac{(2\alpha - 1)K}{2\alpha - 3} \left(e^{-(R_n - y)} - e^{-(\alpha - \frac{1}{2})(R_n - y)}\right)
		&\mbox{if } \alpha > 3/2.
	\end{cases}\\
	&= \Mu{\BallPo{0,y}} \phi_{\H,n}^{(2)}(y)
\end{align*}
and hence
\[
	\Mu{\BallHyp{(0,y)} \cap\Rcal_n[(0,R_n-y)]} = \Mu{\BallPo{0,y}}\left(1 - e^{-(\alpha - \frac{1}{2})(R_n - y)} 
	+ \phi_{\H,n}^{(2)}(y)\right).
\]
Combining this with~\eqref{eq:mu_hyperbolic_ball_part_1} yields the required upper bound.

The lower bound follows by observing that the only difference with the above computations is the change of sign in front of 
\[
	\frac{2 \nu \alpha}{\pi} K e^{\frac{3 y}{2} - R_n}\int_0^{R_n-y} e^{(\frac{3}{2} - \alpha)y^\prime} \dd y^\prime.
\]
\end{proof}

### 2018-10-18 introduction.tex
updated figure 3, adjusted computation lemma 3.2 and added text on scaling k_n

\paragraph{Maximum scaling for $k_n$ is $n^{\frac{1}{2\alpha +1}}$}
%Maybe, it would make sense to incorporate this statement into theorem \ref{thm:asymptotic_local_clustering_hyperbolic}, i.e. that the asymptotic behaviour of $c(k_n)$ is not well-defined for $k_n \gg n^{\frac{1}{2\alpha+1}}$? Resp. claim that a.a.s. the number of vertices of degree $k_n \gg n^{\frac{1}{2\alpha+1}}$ is zero, such that taking the average over these is not well-defined.

>>>>

\paragraph{Maximum scaling for $k_n$ is $n^{\frac{1}{2\alpha +1}}$}
All our results for clustering in the hyperbolic random graph are valid for any sequence $k_n$ such that $k_n = \smallO{n^{\frac{1}{2\alpha + 1}}}$. Although one would like to have results for any sequence $k_n \le n$, it turns out that $n^{1/(2\alpha + 1)}$ is the optimal scaling for which Theorem~\ref{thm:clustering_coefficient_hyperbolic} can be true. To see why this is the case note that by definition of the local clustering function~\eqref{eq:def_local_clustering_general} we have that $c_{\H,n}(k_n) = 0$ if $N_{\H,n}(k_n) = 0$. Hence, it follows by Markov's inequality that for any positive function $f$
\begin{align*}
	\Exp{\left|\frac{c_{\H,n}(k_n)}{f(k_n)} - 1\right|} 
	&\ge \Prob{N_{\H,n}(k_n) = 0} \ge 1 - \Exp{N_{\H,n}(k_n)}.
\end{align*}
We shall later establish (see Lemma [??] \PvdH{Find the reference for this Lemma}) that $\Exp{N_{\H,n}(k_n)} = \bigT{n k_n^{-(2\alpha + 1)}}$. Therefore if $k_n$ is such that $k_n = \omega\left(n^{\frac{1}{2\alpha + 1}}\right)$ we have
\[
	\lim_{n \to \infty} n k_n^{-(2\alpha + 1)} 
	= \lim_{n \to \infty} \left(n^{-\frac{1}{2\alpha + 1}} k_n\right)^{-(2\alpha + 1)} = 0
\]
and hence
\[
	\lim_{n \to \infty} \Exp{\left|\frac{c_{\H,n}(k_n)}{f(k_n)} - 1\right|} 
	\ge \lim_{n \to \infty} 1 - \Exp{N_{\H,n}(k_n)} = \lim_{n \to \infty} 1 - \bigT{n k_n^{-(2\alpha + 1)}} = 1 \ne 0,
\]
for any positive function $f$. This implies that we cannot expect a result like that of Theorem~\ref{thm:clustering_coefficient_hyperbolic} to hold as soon as $k_n = \omega\left(n^{\frac{1}{2\alpha + 1}}\right)$.

%Maybe, it would make sense to incorporate this statement into theorem \ref{thm:asymptotic_local_clustering_hyperbolic}, i.e. that the asymptotic behaviour of $c(k_n)$ is not well-defined for $k_n \gg n^{\frac{1}{2\alpha+1}}$? Resp. claim that a.a.s. the number of vertices of degree $k_n \gg n^{\frac{1}{2\alpha+1}}$ is zero, such that taking the average over these is not well-defined.
